---
title: Effects of Soil Warming Plus Nitrogen Addition on Plant Richness and Diversity
  at Harvard Forest
author: "Roy Moger-Reischer and Katie Beidle, Quantitative Biodiversity, Indiana University"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
geometry: margin=2.54cm
---

## PROJECT OVERVIEW

## ABSTRACT

    Contemporary anthropogenic changes to ecosystem parameters can alter N and C cycling and reduce biodiversity. We examined the effects of warming and N deposition on community composition and soil respiration in a temperate forest. WE FOUND THAT WARMING SIGNIFICANTLY REDUCED CO2 FLUX. TREE SPECIES TURNOVER DID NOT CHANGE SUBSTANTIALLY THROUGH TIME. SPECIES RICHNESS AND DIVERSITY WERE NOT SIGNIFICANTLY DIFFERENT AMONG TREATMENTS. THERE WAS NO INDICATION OF AN INTERACTION EFFECT BETWEEN N AND WARMING MANIPULATIONS. Our analysis indicates ecosystem disturbances do alter element cycling. We suggest that a longer-duration study is needed in order to observe significant effects on tree community dynamics, give their lengthy life histories.
    
## 1)  INTRODUCTION

    Two of the most well-documented outcomes of anthropogenic global change are nitrogen deposition and climatic warming. These changes can alter C and N cycling in forests, and may favor certain species in plant and microbial communities, decreasing biodiversity and leading to species turnover. For example, there is evidence that N addition can reduce CO2 flux across numerous contexts (Bowden et al. 2004; Ramirez et al. 2012). It is known that N deposition can be related to changes in soil fungi community composition, but effects on bacterial communities are less evident (Hesse et al. 2015). Effects on tree communities associated with these microbes may be less facile to observe, given their slower rates of growth and turnover. Similarly, soil warming has been shown to increase soil respiration, while effects on microbial community composition are more ambiguous (Schindlbacher et al. 2011). There is ample research indicating physiological and phonological shifts in individual tree species under a warmed scenario (Xu et al. 2012; Zhao & Liu 2009). There is also evidence that particular species in a community can respond more or less favorably to such change, indicating that long-term warming could alter forest biodiversity (Arend et al. 2011; Thakur et al. 2014).
    In this analysis we evaluate primarily plant community and soil abiotic properties in a temperate forest. Abundances of 40 plant species were censused each year 2006 - 2009. Soil abiotic variables were measured ~monthly during this time. We used R to explore the effects of nitrogen and warming treatments on plant biodiversity and soil carbon and nitrogen dynamics over time.




**Research Question**: What are the short-term effects of soil warming plus N additions on plant richness and diversity at Harvard Forest from 2006-2009?

> ***Hypothesis 1***: Richness and diversity will decrease over time in warming treatments

> ***Hypothesis 2***: Richness and diversity will increase over time with the addition of nitrogen

## 2) SETUP
### Retrieve and Set Working Directory
```{r setup, include=FALSE}
clr = function() {
  ENV = globalenv()
  ll = ls(envir = ENV)
  ll = ll[ll != "clr"]
  rm(list = ll, envir = ENV)
}
getwd() 
setwd("/Users/bhbeidler/GitHub/QB2017_DivPro/Data")
#setwd("C:\\Users\\rmoge\\GitHub\\QB2017_DivPro\\Data")
```
## 3)  DESCRIPTION OF DATA

    The overall objective of this project is to investigate whether soil warming and N fertilization alter plant biodiversity through time (2006-2009).
    In 2006 a group at the Harvard Forest LTER initiated an experiment to assess the interactive effects of these two environmental changes on plant, arthropod, and soil microbe community composition, and soil respiration (Frey 2009). Six replicate plots for each of four treatments (Control; N deposition; heated; N & heated interaction) were established. N deposition plots received 50 kg N/ha/yr. Heated plots were warmed to 5 °C above ambient temperature.
    The data for this project came from the Harvard Forest LTER site. [link](https://portal.lternet.edu/nis/metadataviewer?packageid=knb-lter-hfr.45.24)

### B.Install Packages
This analysis will require several packages. The require() function in R returns TRUE if the package was successfully loaded or FALSE if the package failed to load. This for loop loads each package and installs the package when require() returns FALSE.
```{r results='hide', message=FALSE, warning=FALSE} 
package.list = c('vegan', 'tidyr', 'dplyr', 'codyn', 'ggplot2','betapart','splitstackshape','cowplot', 'MullerPlot', 'RColorBrewer','lubridate', 'reshape2', 'lubridate','wesanderson','TTR', 'xtable', 'multcomp', 'pander', 'png', 'grid', 'tseries', 'nlme', 'forecast', 'lsmeans', 'devtools')

for (package in package.list) {
if (!require(package, character.only = TRUE, quietly = TRUE)) {
install.packages(package, repos='http://cran.us.r-project.org')
library(package, character.only = TRUE) }
}
```
## 2) LOADING DATA

### A. Description of Data Set
[Explain what is contained in the different data sheets...]
1. 'plant'
2. 'nmin'
3. 'resp'

```{r}
setwd("/Users/bhbeidler/GitHub/QB2017_DivPro/Data")
plant = read.csv("./HF_plants_treat.csv")
nmin = read.csv("./HF_nmin.csv", stringsAsFactors = FALSE)
resp = read.csv("./HF_soilresp.csv", stringsAsFactors = FALSE)


#plant = read.csv("C:\\Users\\rmoge\\GitHub\\QB2017_DivPro\\Data\\HF_plants_treat.csv")
#nmin = read.csv("C:\\Users\\rmoge\\GitHub\\QB2017_DivPro\\Data\\HF_nmin.csv", stringsAsFactors = FALSE)
#resp = read.csv("C:\\Users\\rmoge\\GitHub\\QB2017_DivPro\\Data\\HF_soilresp.csv", stringsAsFactors = FALSE)

```
### B. Data Wrangling
Subsetting the data by year and producing site by species matrices. 
```{r}
# Making the Site by Species Matrix for the plant data set
plant_sbys = plant[ ,6:43]

# Subsetting the data into the different years
plant_06 = (filter(plant, year == 2006))
plant_07 = (filter(plant, year == 2007))
plant_08 = (filter(plant, year == 2008))
plant_09 = (filter(plant, year == 2009))

# Separating out the treatments from the site by species matrices 
plant_06_sbys = plant_06[ ,6:43]
plant_07_sbys = plant_07[ ,6:43]
plant_08_sbys = plant_08[ ,6:43]
plant_09_sbys = plant_09[ ,6:43]
```
###  BIODIVERSITY DATA VISUALIZATION & PLOTS
[plots that show richness for each site/ abundance for a particular species]
Visualizing Beta Diversity
```{r}
# Visualizing the Plant data set
str(plant, max.level = 1)

# Calculate observed richness from time-by-species matrix
p_richness = as.data.frame(rowSums(plant[,-c(1:5)] > 0)) 
# Create data frame with experimental design and richness data
p_rich.all = data.frame(plant[,1:5,], p_richness)
# Rename column
names(p_rich.all)[6] = "richness"

# avg. richness per group
# stand. dev. per group
# num. obs. per group
# calc. std. err. mean.
p_rich.treat.plot = group_by(p_rich.all, treatment, year) %>%
  summarise(mean = mean(richness), sd = sd(richness),n = n(),sem = sd/sqrt(n))

p_rich.plot = ggplot(p_rich.treat.plot, aes(x = year, y = mean, color = as.factor(treatment))) + 
               geom_point(size = 2, show.legend = T) +
               geom_line(size = 0.75) +
               geom_errorbar(aes(ymin = mean - sem, ymax = mean + sem), width = .1) +
               scale_color_manual(values = c("#999999", "#E69F00", "#000000", "#56B4E9"),name="Soil Treatment",labels=c("Unheated-Unfertilized", "Heated-Unfertilized", "Heated- Fertilized", "Unheated-Fertilized")) +
               xlim(2006, 2009) + xlab("Year") + ylab("Richness") +
               theme_classic(base_size=15)+
               theme(axis.line.x = element_line(color = "black"), axis.line.y = element_line(color ="black"))
plot(p_rich.plot)

# Calculate diveristy from time-by-species matrix
p_diversity = diversity(plant[,-c(1:5)], "shannon")

# Create data frame with experimental design and diversity data
p_div.all = data.frame(plant[,1:5,], p_diversity)
# Rename column
names(p_div.all)[6] = "diversity"

p_div.treat.plot = group_by(p_div.all, treatment, year) %>%
  summarise(mean = mean(diversity), sd = sd(diversity),n = n(),sem = sd/sqrt(n))

p_div.plot = ggplot(p_div.treat.plot, aes(x = year, y = mean, color = as.factor(treatment))) + 
               geom_point(size = 2, show.legend = T) +
               geom_line(size = 0.75) +
               geom_errorbar(aes(ymin = mean - sem, ymax = mean + sem), width = .1) +
               scale_color_manual(values = c("#999999", "#E69F00", "#000000", "#56B4E9"),name="Soil Treatment",labels=c("Unheated-Unfertilized", "Heated-Unfertilized", "Heated- Fertilized", "Unheated-Fertilized")) +
               xlim(2006, 2009) + xlab("Year") + ylab("Shannon's Diversity Index")+
               theme_classic(base_size=15)+
               theme(axis.line.x = element_line(color = "black"),axis.line.y = element_line(color = "black"))
plot(p_div.plot)
```
### BIODIVERSITY HYPOTHESIS TESTING
```{r}
p_rich.rm = lme(richness ~ year + heat_treat + N_treat * heat_treat * N_treat * year, random = ~ 1 | plot, correlation = corAR1(form = ~ 1 | plot),data = p_rich.all) 
summary(p_rich.rm) # Obtain F-test
anova(p_rich.rm)
# Make cleaner ANOVA table
set.caption("RMANOVA for Plant Richness Data") 
pander(anova(p_rich.rm))
 
# Perform an RM-ANOVA and construct a F-test using the AR(1)
p_div.rm = lme(diversity ~ year + heat_treat + N_treat * heat_treat * N_treat * year, random = ~ 1 | plot, correlation = corAR1(form = ~ 1 | plot),data = p_div.all) # Look at detailed output
summary(p_div.rm) # Obtain F-test
anova(p_div.rm)
set.caption("RMANOVA for Plant Diversity Data") 
pander(anova(p_div.rm))
```
### TEMPORAL BIODIVERSITY
### A. Turnover
```{r}
# Making the wide site by species format into a long format- to get abundances for each species 
plant$row_id = 1:nrow(plant)
p.id = dplyr::select(plant, row_id, year, treatment)
plant.m = as.matrix(plant_sbys)
m1 = setNames(melt(plant.m), c('row_id','species','count'))
m2 = dplyr::filter(m1, count > 0)
m3 = expandRows(m2, "count")
plant_long = left_join(p.id,m3, by = "row_id")
p_long = dplyr::select(plant_long, -row_id)

# Calculate species abundances for each taxonomic group 
plant.sp.abunds = p_long %>% 
                     group_by(year,treatment) %>% 
                     count(species)

# Calculate total turnover
plant.total = turnover(df = plant.sp.abunds, time.var = "year",
                            species.var = "species",
                            abundance.var = "n",
                            replicate.var = "treatment",
                            metric = "total")
# Calculate species gained
plant.appearance = turnover(df = plant.sp.abunds, time.var = "year",
                            species.var = "species",
                            abundance.var = "n",
                            replicate.var = "treatment",
                            metric = "appearance")
# Calculate species lost
plant.disappearance = turnover(df = plant.sp.abunds, time.var = "year",
                            species.var = "species",
                            abundance.var = "n",
                            replicate.var = "treatment",
                            metric = "disappearance")

plant.turnover = full_join(plant.total, plant.disappearance) %>% 
                  full_join(plant.appearance)

plant.turnover = gather(plant.turnover, key = metric, value = turnover, total, appearance, disappearance)
View(plant.turnover)
# 3. Visualize turnover within each group

plant.turn.plot = ggplot(plant.turnover, aes(x = year, y = turnover, color = metric)) +
  geom_line(size = 1, show.legend = T) + facet_wrap(~treatment, ncol = 1) +
  xlim(2007, 2009) +
  xlab("Year") +
  ylab("Turnover") +
  theme(legend.position = "bottom") +
  scale_color_grey()
plot(plant.turn.plot)
# Low turnover is indicative of a stable community and high turnover is indicative of a dynamic community

# Abundance Plots
p_abun.plot = ggplot(plant.sp.abunds, aes(year, n, color= as.factor(treatment))) + 
              geom_point() +
              geom_line(aes(color = as.factor(treatment))) +
              facet_wrap(~species, nrow = 8) +
              scale_color_manual(values = c("#999999", "#E69F00", "#000000", "#56B4E9"),name="Soil Treatment",labels=c("Unheated-Unfertilized", "Heated-Unfertilized", "Heated- Fertilized", "Unheated-Fertilized")) + 
              ylab("# of  Stems") + 
              theme_bw() +
              theme(axis.text.x = element_text(angle=70, vjust=0.5))

# Abundances at time point 1
p_abun.06 = dplyr::filter(plant.sp.abunds, year == "2006")
p_abun.plot.06 = ggplot(p_abun.06 , aes(x = species, y = n)) + 
                  geom_bar(stat = "identity") + 
                  facet_wrap(~treatment, ncol = 1) +
                  ylab("# of  Stems") +
                  theme_bw() +
                  theme(axis.text.x = element_text(angle=70, vjust=0.5)) 

# Abundances at time point 2
p_abun.09 = dplyr::filter(plant.sp.abunds, year == "2009")
p_abun.plot.09 = ggplot(p_abun.09 , aes(x = species, y = n)) + 
                  geom_bar(stat = "identity") + 
                  facet_wrap(~treatment, ncol = 1) +
                  ylab("# of  Stems") +
                  theme_bw() +
                  theme(axis.text.x = element_text(angle=70, vjust=0.5)) 

```
### B. Rank Shift
```{r}
# Calculate species abundances for each taxonomic group 
plant.sp.abunds = p_long %>% 
                     group_by(year,treatment) %>% 
                     count(species)

# Calculate MRS
plant.rankshift = rank_shift(df = as.data.frame(plant.sp.abunds), 
                              time.var = "year",
                              species.var = "species",
                              abundance.var = "n",
                              replicate.var = "treatment")

plant.rankshift$year = as.numeric(substr(plant.rankshift$year_pair, 6, 9))

# Create ggplot
rankshift.plot =  ggplot(plant.rankshift, aes(x = year, y = MRS, color = treatment)) + geom_line(size = 1) +
xlim(2007, 2009) +
xlab("Year") +
ylab("Mean Rank Shift") + scale_color_grey()
plot(rankshift.plot)
portal.rankshift %>% 
  group_by (plot_type) %>% 
  summarise(mean = mean(MRS), cv = sd(MRS)/mean)

# Rank Clock
ag_plant.dat = aggregate(n ~ species * year * treatment, 
                    data = subset(plant.sp.abunds, 
                                    species == "denpun" |
                                    species == "maican" | 
                                    species == "vaccspp"| 
                                    species == "gaupro"), FUN = mean)

rclock_plot = ggplot(ag_plant.dat, aes(year, n, color = species)) + 
        geom_line(size = 1) + coord_polar() + theme_bw() +
        facet_wrap(~treatment) +
        ylab("# of Stems") +
        scale_color_manual(values = c("#999999", "#E69F00", "#000000", "#56B4E9")) +
        ggtitle("Dominant species abundances \n for the different soil treatments, Harvard Forest \n")
plot(rclock_plot)
```

### SOIL MEASURMENTS 
### A. Plots
```{r}
# Soil respiration
resp = resp %>%
    filter(year %in% c( "2006","2007", "2008", "2009"))
resp$date = as.Date(resp$date, format = "%m/%d/%Y")
resp$doy = yday(resp$date)

resp = na.exclude(resp) 

resp_yr.plot = qplot(doy, co2flux, data = resp, geom="smooth") +
         stat_smooth(fill="azure3", colour="mediumpurple4", size=1, alpha = 0.2) + 
         theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
            panel.background = element_blank(), axis.line =   element_line(colour = "black")) + 
        facet_wrap(~year, ncol = 1)
        xlab("Day of the Year") +
        ylab("CO2 Flux") 
resp_yr.plot + theme(text = element_text(size=18), plot.title = element_text(lineheight=.8, face="bold")) + ggtitle("Soil Respiration 2006-2009")
plot(resp_yr.plot)

resp_trt.plot = qplot(doy, co2flux, data = resp, colour = as.factor(trt), geom="smooth", se=FALSE) +
      stat_smooth(method ="auto",se = FALSE, formula = y ~ x, size = 1) +
      facet_wrap(~year, ncol = 1) +
      theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
            panel.background = element_blank(), axis.line = element_line(colour = "black")) 
resp_trt.plot + scale_colour_brewer(type="seq", palette=3) + xlab("Day of the Year") +ylab("CO2 Flux)") + theme(text = element_text(size=18))
plot(resp_trt.plot)

# N mineralization
nmin = nmin %>%
    filter(year %in% c("2006","2007", "2008", "2009"))
nmin$date = as.Date(nmin$date, format = "%m/%d/%Y")
nmin$doy = yday(nmin$date)
nmin = na.remove(nmin)

n03_yr.plot = qplot(doy, netno3, data = nmin, geom="smooth") +
         stat_smooth(fill="azure3", colour="mediumpurple4", size=1, alpha = 0.2) + 
         theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
            panel.background = element_blank(), axis.line =   element_line(colour = "black")) + 
        facet_wrap(hor~year)
        xlab("Day of the Year") +
        ylab("Net Nitrate") 
nmin_yr.plot + theme(text = element_text(size=18), plot.title = element_text(lineheight=.8, face="bold")) + ggtitle("Soil NO3 2006-2009")
plot(n03_yr.plot)

nh4_yr.plot = qplot(doy, netnh4, data = nmin, geom="smooth") +
         stat_smooth(fill="azure3", colour="mediumpurple4", size=1, alpha = 0.2) + 
         theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
            panel.background = element_blank(), axis.line =   element_line(colour = "black")) + 
        facet_wrap(hor~year)
        xlab("Day of the Year") +
        ylab("Net Ammonium") 
nh4_yr.plot + theme(text = element_text(size=18), plot.title = element_text(lineheight=.8, face="bold")) + ggtitle("Soil NO3 2006-2009")
plot(nh4_yr.plot)

no3_trt.plot = qplot(doy, netnh4, data = nmin, colour = as.factor(trt), geom="smooth", se=FALSE) +
      stat_smooth(method ="auto",se = FALSE, formula = y ~ x, size = 1) +
      facet_wrap(hor~year) +
      theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
            panel.background = element_blank(), axis.line = element_line(colour = "black")) 
resp_trt.plot + scale_colour_brewer(type="seq", palette=3) + xlab("Day of the Year") +ylab("Total NO3)") + theme(text = element_text(size=18))
plot(no3_trt.plot)

nh4_trt.plot = qplot(doy, netnh4, data = nmin, colour = as.factor(trt), geom="smooth", se=FALSE) +
      stat_smooth(method ="auto",se = FALSE, formula = y ~ x, size = 1) +
      facet_wrap(hor~year) +
      theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
            panel.background = element_blank(), axis.line = element_line(colour = "black")) 
nh4_trt.plot + scale_colour_brewer(type="seq", palette=3) + xlab("Day of the Year") +ylab("Total NH4)") + theme(text = element_text(size=18))
plot(nh4_trt.plot)

```
### B. Hypothesis Testing
```{r} 
plant.treat.plot = group_by(resp, trt, year) %>%
  summarise(mean = mean(co2flux), sd = sd(co2flux),n = n(),sem = sd/sqrt(n)) 

# Trying different variance - covariance structures 
# AR(1) covariance structure
soil.rm1 = lme(co2flux ~ year + heat_treat + N_treat * heat_treat * N_treat * year , random = ~ 1 | plot, correlation = corAR1(form = ~ 1 | plot),data = resp) 
summary(soil.rm1) # Obtain F-test
anova(soil.rm1)

# corARMA autoregressive moving average proccess 
soil.rm2 = lme(co2flux ~ year + heat_treat + N_treat * heat_treat * N_treat * year , random = ~ 1 | plot, correlation = corCompSymm(form = ~ 1 | plot),data = resp) #
summary(soil.rm2) # Obtain F-test
anova(soil.rm2)

# cirCAR1 cnt AR process
soil.rm3 = lme(co2flux ~ year + heat_treat + N_treat * heat_treat * N_treat * year , random = ~ 1 | plot, correlation = corCAR1(form = ~ 1 | plot),data = resp) 
summary(soil.rm3) # Obtain F-test
anova(soil.rm3)

# corARMA autoregressive moving average process, with arbitrary components
soil.rm4 = lme(co2flux ~ year + heat_treat + N_treat * heat_treat * N_treat * year , random = ~ 1 | plot, correlation = corARMA(form = ~ 1 | plot, p = 1, q = 1),data = resp) 
summary(soil.rm4) # Obtain F-test
anova(soil.rm4)

#corExp exponential spatial correlation.
soil.rm5 = lme(co2flux ~ year + heat_treat + N_treat * heat_treat * N_treat * year , random = ~ 1 | plot, correlation = corExp(form = ~ 1 | plot),data = resp) 
summary(soil.rm5) # Obtain F-test
anova(soil.rm5)

# CorGaus Gaussian spatial correlation.
soil.rm6 = lme(co2flux ~ year + heat_treat + N_treat * heat_treat * N_treat * year , random = ~ 1 | plot, correlation = corGaus(form = ~ 1 | plot),data = resp) 
summary(soil.rm6) # Obtain F-test
anova(soil.rm6)

# corLin	linear spatial correlation.
soil.rm7 = lme(co2flux ~ year + heat_treat + N_treat * heat_treat * N_treat * year , random = ~ 1 | plot, correlation = corLin(form = ~ 1 | plot),data = resp) #
summary(soil.rm7) # Obtain F-test
anova(soil.rm7)

# corRatio	Rational quadratics spatial correlation.
soil.rm8 = lme(co2flux ~ year + heat_treat + N_treat * heat_treat * N_treat * year , random = ~ 1 | plot, correlation = corRatio(form = ~ 1 | plot),data = resp) #
summary(soil.rm8) # Obtain F-test
anova(soil.rm8)

# corRatioSpher	spherical spatial correlation
soil.rm9 = lme(co2flux ~ year + heat_treat + N_treat * heat_treat * N_treat * year , random = ~ 1 | plot, correlation = corSpher(form = ~ 1 | plot),data = resp) #
summary(soil.rm9) # Obtain F-test
anova(soil.rm9)

# corSymm general correlation matrix, with no additional structure
soil.rm10 = lme(co2flux ~ year + heat_treat + N_treat * heat_treat * N_treat * year , random = ~ 1 | plot, correlation = corSpher(form = ~ 1 | plot),data = resp) #
summary(soil.rm10) # Obtain F-test
anova(soil.rm10)

# Compare the AICs
AIC(soil.rm1, soil.rm2, soil.rm3,soil.rm4, soil.rm5, soil.rm6,soil.rm7,soil.rm8,soil.rm9,soil.rm10)

# The unstructured covariance matrix has the lowest AIC score ??
```

## 6)  DISCUSSION AND CONCLUSION

HELP A TROLL HELP A TROLL HELP A TROLL HELP A TROLL HELP A TROLL HELP A TROLL HELP A TROLL HELP A TROLL HELP A TROLL HELP A TROLL HELP A TROLL HELP A TROLL HELP A TROLL HELP A TROLL HELP A TROLL HELP A TROLL HELP A TROLL HELP A TROLL HELP A TROLL HELP A TROLL HELP A TROLL HELP A TROLL HELP A TROLL HELP A TROLL HELP A TROLL HELP A TROLL HELP A TROLL HELP A TROLL HELP A TROLL HELP A TROLL HELP A TROLL HELP A TROLL HELP A TROLL HELP A TROLL HELP A TROLL HELP A TROLL HELP A TROLL HELP A TROLL HELP A TROLL HELP A TROLL HELP A TROLL HELP A TROLL HELP A TROLL HELP A TROLL HELP A TROLL HELP A TROLL HELP A TROLL HELP A TROLL HELP A TROLL HELP A TROLL 

## 7)  REFERENCES

Arend, M., Kuster, T., Günthardt-Goerg, M.S. and Dobbertin, M., 2011. Provenance-specific growth responses to drought and air warming in three European oak species (Quercus robur, Q. petraea and Q. pubescens). Tree Physiology, 31(3), pp.287-297.

Bowden, R.D., Davidson, E., Savage, K., Arabia, C. and Steudler, P., 2004. Chronic nitrogen additions reduce total soil respiration and microbial respiration in temperate forest soils at the Harvard Forest. Forest Ecology and Management, 196(1), pp.43-56.

Frey, Serita (2009): Soil warming plus nitrogen addition expermient at Harvard Forest since 2006. Long Term Ecological Research Network. [link]http://dx.doi.org/10.6073/pasta/08505f370bbcf98b7237434e41c78bc8


Hesse, C.N., Mueller, R.C., Vuyisich, M., Gallegos-Graves, L.V., Gleasner, C.D., Zak, D.R. and Kuske, C.R., 2015. Forest floor community metatranscriptomes identify fungal and bacterial responses to N deposition in two maple forests. Frontiers in microbiology, 6, p.337.

Ramirez, K.S., Craine, J.M. and Fierer, N., 2012. Consistent effects of nitrogen amendments on soil microbial communities and processes across biomes. Global Change Biology, 18(6), pp.1918-1927

Schindlbacher, A., Rodler, A., Kuffner, M., Kitzler, B., Sessitsch, A. and Zechmeister-Boltenstern, S., 2011. Experimental warming effects on the microbial community of a temperate mountain forest soil. Soil Biology and Biochemistry, 43(7), pp.1417-1425.

Thakur, M.P., Reich, P.B., Eddy, W.C., Stefanski, A., Rich, R., Hobbie, S.E. and Eisenhauer, N., 2014. Some plants like it warmer: Increased growth of three selected invasive plant species in soils with a history of experimental warming. Pedobiologia, 57(1), pp.57-60.

Xu, Z., Hu, T. and Zhang, Y., 2012. Effects of experimental warming on phenology, growth and gas exchange of treeline birch (Betula utilis) saplings, Eastern Tibetan Plateau, China. European Journal of Forest Research, 131(3), pp.811-819.

Zhao, C. and Liu, Q., 2009. Growth and physiological responses of Picea asperata seedlings to elevated temperature and to nitrogen fertilization. Acta physiologiae plantarum, 31(1), p.163.