---
title: "Responses of Plants and Ants to Soil Warming and N Fertilization at Harvard Forest"
author: "Roy Moger-Reischer and Katie Beidler"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
---
## OVERVIEW
Insert Description of Project Background (KVB)
Questions we are exploring
a. Does diversity vary among sites? Does this correspond to treatment type?
b. Is treatment type a significant predictor of site dissimilarity?

## 1) Setup 
### Retrieve and sSet your Working Directory
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
clr = function() {
  ENV = globalenv()
  ll = ls(envir = ENV)
  ll = ll[ll != "clr"]
  rm(list = ll, envir = ENV)
}
getwd() 
#setwd("/Users/bhbeidler/GitHub/QB2017_DivPro")
#setwd("C:\\Users\\rmoge\\GitHub\\QB2017_DivPro")
```
### Install Packages
This analysis will require several packages. The require() function in R returns TRUE if the package was successfully loaded or FALSE if the package failed to load. This for loop loads each package and installs the package when require() returns FALSE.
```{r results='hide', message=FALSE, warning=FALSE} 

package.list = c('vegan', 'tidyr', 'dplyr', 'codyn', 'ggplot2', 'cluster',
'cowplot', 'MullerPlot', 'RColorBrewer', 'reshape2', 'lubridate', 'BiodiversityR',
'TTR', 'xtable', 'multcomp', 'pander', 'png', 'grid', 'tseries', 'nlme', 'forecast', 'lsmeans', 'devtools', 'CatterPlots')
for (package in package.list) {
if (!require(package, character.only = TRUE, quietly = TRUE)) {
install.packages(package, repos='http://cran.us.r-project.org')
library(package, character.only = TRUE) }
}
```
## 2) LOADING DATA
### Load datasets
The R code chunk below, does the following: Loads in the plant dataset (explanation), ant dataset (explanation)...Describe the different datasets here
```{r}
#plant = read.csv("./Data/HF_plants.csv")
#ant = read.csv("./Data/HF_ants.csv")
#nmin = read.csv("./Data/HF_nmin.csv")
#resp = read.csv("./Data/HF_soilresp.csv")
#temp = read.csv("./Data/HF_soiltemp.csv")

plant = read.csv("C:\\Users\\rmoge\\GitHub\\QB2017_DivPro\\Data\\HF_plants.csv")
ant = read.csv("C:\\Users\\rmoge\\GitHub\\QB2017_DivPro\\Data\\HF_ants.csv")
nmin = read.csv("C:\\Users\\rmoge\\GitHub\\QB2017_DivPro\\Data\\HF_nmin.csv")
resp = read.csv("C:\\Users\\rmoge\\GitHub\\QB2017_DivPro\\Data\\HF_soilresp.csv")
temp = read.csv("C:\\Users\\rmoge\\GitHub\\QB2017_DivPro\\Data\\HF_soiltemp.csv")
cco2<-read.csv("C:\\Users\\rmoge\\GitHub\\QB2017_DivPro\\Data\\Cflux.csv")
hco2<-read.csv("C:\\Users\\rmoge\\GitHub\\QB2017_DivPro\\Data\\Hflux.csv")
hnco2<-read.csv("C:\\Users\\rmoge\\GitHub\\QB2017_DivPro\\Data\\HNflux.csv")
cco2<-read.csv("C:\\Users\\rmoge\\GitHub\\QB2017_DivPro\\Data\\Cflux.csv")
```
## 2) Data Wrangling
```{r}
# Subsetting the data into the different years
plant_06 = (filter(plant, year == 2006))
plant_07 = (filter(plant, year == 2007))
plant_08 = (filter(plant, year == 2008))
plant_09 = (filter(plant, year == 2009))

ant_06 = filter(ant, year == 2006)
ant_07 = filter(ant, year == 2007)
ant_08 = filter(ant, year == 2008)
ant_09 = filter(ant, year == 2009)

# Separating out the treatments from the site by species matrices 
plant_06_sbys = plant_06[ ,4:43]
plant_07_sbys = plant_07[ ,4:43]
plant_08_sbys = plant_08[ ,4:43]
plant_09_sbys = plant_09[ ,4:43]

ant_06_sbys = ant_06[ ,4:18]
ant_07_sbys = ant_07[ ,4:18]
ant_08_sbys = ant_08[ ,4:18]
ant_09_sbys = ant_09[ ,4:18]
```
## 2) Describe how richness and biodiversity estimates vary among sites.
### Boxlots of Richness/ diversity for the different treatments/ years
### Rank abundance curves
```{r}
# Richness Function
S.obs <- function(x =" "){
  rowSums(x>0) *1 
}
```
## 3) Testing H1: Plant beta Diversity differs among soil treatments using a PERMANOVA
```{r}
# Ordination
# construct a resemblance matrix based on Bray-Curtis Distance ("plant_2006.db") for 2006
plant_06.db = vegdist(plant_06_sbys, method = "bray", upper = TRUE, diag = TRUE)
# Perform a Principal Coordinates Analysis to visualize beta-diversity
plant_06.pcoa = cmdscale(plant_06.db, eig = TRUE, k = 3)

# Calculate the variation explained by the first three axes in your ordination
explainvar1.p = round(plant_06.pcoa$eig[1] / sum(plant_06.pcoa$eig), 3) * 100
explainvar2.p = round(plant_06.pcoa$eig[2] / sum(plant_06.pcoa$eig), 3) * 100
explainvar3.p = round(plant_06.pcoa$eig[3] / sum(plant_06.pcoa$eig), 3) * 100
sum.eig = sum(explainvar1.p, explainvar2.p, explainvar3.p)

# plot the PCoA ordination
# Define Plot Parameters
par(mar = c(5, 5, 1, 2) + 0.1)
# Initiate Plot
plot(plant_06.pcoa$points[ ,1], plant_06.pcoa$points[ ,2], 
xlab = paste("PCoA 1 (", explainvar1.p, "%)", sep = ""),
ylab = paste("PCoA 2 (", explainvar2.p, "%)", sep = ""),
main = "Ordination for 2006",
pch = 16, cex = 2.0, type = "n", cex.lab = 1.5, cex.axis = 1.2, axes = FALSE)
# Add Axes
axis(side = 1, labels = T, lwd.ticks = 2, cex.axis = 1.2, las = 1)
axis(side = 2, labels = T, lwd.ticks = 2, cex.axis = 1.2, las = 1)
abline(h = 0, v = 0, lty = 3)
box(lwd = 2)

# Add Points & Labels
points(plant_06.pcoa$points[ ,1], plant_06.pcoa$points[ ,2],
pch = 16, cex = 3, bg = "gray", col = "gray")
text(plant_06.pcoa$points[ ,1], plant_06.pcoa$points[ ,2],
labels = row.names(plant_06.pcoa$points))

# Identify influential species 
# First we calculate the relative abundances of each species at each site
plant_06REL = plant_06_sbys 
for(i in 1:nrow(plant_06_sbys )){
plant_06REL[i, ] = plant_06_sbys [i, ] / sum(plant_06_sbys [i, ])
}
# Now, we use this information to calculate and add species scores
plant_06.pcoa= add.spec.scores(plant_06.pcoa,plant_06REL,method = "pcoa.scores")
text(plant_06.pcoa$cproj[ ,1], plant_06.pcoa$cproj[ ,2],
labels = row.names(plant_06.pcoa$cproj), col = "black")

# Hypothesis Testing
# Soil treatment (either warming or N fertilization) influenced tree beta diversity during 2006. 

# Multivariate Procedures for Categorical designs 

# There are 40 tree species. For each year and each treatment there are 6 sites. Treatments include: 1=control, 2=warm, 3=nitrogen, 4=warm/nitrogen
# Create "Factors" vector for categorizing soil treatments
soil_treat = as.factor(plant_06[ ,3])

# Run PERMANOVA with adonis function
P_PERMA.06 = adonis(plant_06_sbys ~ soil_treat, method = "bray", permutations = 999)
P_PERMA.07 = adonis(plant_07_sbys ~ soil_treat, method = "bray", permutations = 999)
P_PERMA.08 = adonis(plant_08_sbys ~ soil_treat, method = "bray", permutations = 999)
P_PERMA.09 = adonis(plant_09_sbys ~ soil_treat, method = "bray", permutations = 999)

P_PERMA.06 
P_PERMA.07
P_PERMA.08 
P_PERMA.09
```
## 4) Testing H2: Ant beta Diversity differs among soil treatments using a PERMANOVA
```{r}
# Removing columns 
ant_06_sbys = ant_06_sbys[, !apply(ant_06_sbys == 0, 2, all)]
ant_06_sbys = rowSums(ant_06_sbys == 0) != ncol(ant_06_sbys)

ant_07_sbys = ant_07_sbys[, !apply(ant_07_sbys == 0, 2, all)]
ant_08_sbys = ant_08_sbys[, !apply(ant_08_sbys == 0, 2, all)]
ant_09_sbys = ant_09_sbys[, !apply(ant_09_sbys == 0, 2, all)]

# Ordination
# construct a resemblance matrix based on Bray-Curtis Distance ("plant_2006.db") for 2006
ant_06.db = vegdist(ant_06_sbys, method = "bray")
# Perform a Principal Coordinates Analysis to visualize beta-diversity
ant_06.pcoa = cmdscale(ant_06.db, eig = TRUE, k = 3)

# Calculate the variation explained by the first three axes in your ordination
explainvar1.a = round(ant_06.pcoa$eig[1] / sum(ant_06.pcoa$eig), 3) * 100
explainvar2.a = round(ant_06.pcoa$eig[2] / sum(ant_06.pcoa$eig), 3) * 100
explainvar3.a = round(ant_06.pcoa$eig[3] / sum(ant_06.pcoa$eig), 3) * 100
sum.eig = sum(explainvar1.a, explainvar2.a, explainvar3.a)

# plot the PCoA ordination
# Define Plot Parameters
par(mar = c(5, 5, 1, 2) + 0.1)
# Initiate Plot
plot(ant_06.pcoa$points[ ,1], ant_06.pcoa$points[ ,2], 
xlab = paste("PCoA 1 (", explainvar1.a, "%)", sep = ""),
ylab = paste("PCoA 2 (", explainvar2.a, "%)", sep = ""),
main = "Ant Ordination for 2006",
pch = 16, cex = 2.0, type = "n", cex.lab = 1.5, cex.axis = 1.2, axes = FALSE)
# Add Axes
axis(side = 1, labels = T, lwd.ticks = 2, cex.axis = 1.2, las = 1)
axis(side = 2, labels = T, lwd.ticks = 2, cex.axis = 1.2, las = 1)
abline(h = 0, v = 0, lty = 3)
box(lwd = 2)

# Add Points & Labels
points(plant_06.pcoa$points[ ,1], ant_06.pcoa$points[ ,2],
pch = 16, cex = 3, bg = "gray", col = "gray")
text(ant_06.pcoa$points[ ,1], ant_06.pcoa$points[ ,2],
labels = row.names(ant_06.pcoa$points))

# Identify influential species 
# First we calculate the relative abundances of each species at each site
ant_06REL = ant_06_sbys 
for(i in 1:nrow(ant_06_sbys )){
ant_06REL[i, ] = ant_06_sbys [i, ] / sum(ant_06_sbys [i, ])
}
# Now, we use this information to calculate and add species scores
ant_06.pcoa= add.spec.scores(ant_06.pcoa,ant_06REL,method = "pcoa.scores")
text(ant_06.pcoa$cproj[ ,1], ant_06.pcoa$cproj[ ,2],
labels = row.names(ant_06.pcoa$cproj), col = "black")

# Hypothesis Testing
# H2: Soil treatment (either warming or N fertilization) influenced beta diversity of ants for the final time period 2009. 

# Multivariate Procedures for Categorical designs 

# There are 18 ant species. For each year and each treatment there are 6 sites. Treatments include: 1=control, 2=warm, 3=nitrogen, 4=warm/nitrogen
# Create "Factors" vector for categorizing soil treatments
soil_treat = as.factor(ant_06[ ,3])

# Run PERMANOVA with adonis function
A_PERMA.06 = adonis(ant_06_sbys ~ soil_treat, method = "bray", permutations = 999)
A_PERMA.07 = adonis(ant_07_sbys ~ soil_treat, method = "bray", permutations = 999)
A_PERMA.08 = adonis(ant_08_sbys ~ soil_treat, method = "bray", permutations = 999)
A_PERMA.09 = adonis(ant_09_sbys ~ soil_treat, method = "bray", permutations = 999)

A_PERMA.06 
A_PERMA.07
A_PERMA.08 
A_PERMA.09
```
## 5) Testing H3: plant richness (diversity) differs among years and soil treatments 
```{r}
# Calculate observed richness from time-by-species matrix
p_richness = as.data.frame(rowSums(plant[,-c(1:3)] > 0)) 
# Create data frame with experimental design and richness data
p_rich.all = data.frame(plant[,1:3,], richness)
# Rename column
names(p_rich.all)[4] = "richness"

p_rich.treat.plot = group_by(p_rich.all, treatment, year) %>%
  summarise(mean = mean(richness), sd = sd(richness),n = n(),sem = sd/sqrt(n))
# avg. richness per group
# stand. dev. per group
# num. obs. per group
# calc. std. err. mean.
p_rich.plot = ggplot(p_rich.treat.plot, aes(x = year, y = mean, color = treatment)) + 
                  geom_point(size = 1, show.legend = T) +
                  geom_errorbar(aes(ymin = mean - sem, ymax = mean + sem), width = .1) +  
                  xlim(2006, 2009) + xlab("Year") + ylab("Richness")
plot(p_rich.plot)
# 3. Perform an RM-ANOVA and construct a F-test using the AR(1)
p_rich.rm = lme(richness ~ treatment * year, random = ~ 1 | treatment, correlation = corAR1(form = ~ 1 | treatment),data = p_rich.all) # Look at detailed output
summary(p_rich.rm) # Obtain F-test
anova(p_rich.rm)
# Make cleaner ANOVA table
set.caption("RMANOVA for Plant Data") 
pander(anova(p_rich.rm))
# Use `lsmeans` package for time-corrected marginal means
lsmeans(p_rich.rm, ~ treatment)
```
## 6) Testing H4: ant richness (diversity) differs among years and soil treatments
```{r}

```
## 7) Testing H5: Soil respiration is positively correlated with richness
```{r}

plants.sortedT<-arrange(plant, treatment)
plants.sortedY<-arrange(plants.sortedT,year)
plants.sortedP<-arrange(plants.sortedY,plot)
my2009plants.sortedY<-filter(plants.sortedY, year==2009, plot!=16 & plot!=12 & plot!=19)
#dim(my2009plants.sortedY)
sbs2009plants.sortedYt<-my2009plants.sortedY[4:43]
sbs2009plants.sortedYtt<-as.data.frame(sbs2009plants.sortedYt)

my2008plants.sortedY<-filter(plants.sortedY, year==2008)
#dim(my2008plants.sortedY)
sbs2008plants.sortedYt<-my2008plants.sortedY[4:43]
sbs2008plants.sortedYtt<-as.data.frame(sbs2008plants.sortedYt)
my2007plants.sortedY<-filter(plants.sortedY, year==2007)
#dim(my2007plants.sortedY)
sbs2007plants.sortedYt<-my2007plants.sortedY[4:43]
sbs2007plants.sortedYtt<-as.data.frame(sbs2007plants.sortedYt)
my2006plants.sortedY<-filter(plants.sortedY, year==2006)
#dim(my2006plants.sortedY)
sbs2006plants.sortedYt<-my2006plants.sortedY[4:43]
sbs2006plants.sortedYtt<-as.data.frame(sbs2006plants.sortedYt)

soilenv<-read.csv("C:\\Users\\rmoge\\GitHub\\QB2017_DivPro\\Data\\HF_soilresp.csv")
mysoilenv_holes<-na.omit(soilenv,year==2009,month==1)
my2009soilenv_1<-filter(soilenv,year==2009,month==1,plot!=12 & plot!=16 & plot!=19)
co2_1<-c(my2009soilenv_1[8])
co2_1<-
print(co2_1)
tester<-c(S.obs(sbs2009plants.sortedYtt))
sbs2009_1.sobs<-c(tester,tester)
toyr<-cbind(data.frame(sbs2009_1.sobs, co2_1))
lmfit<-lm(co2flux~sbs2009_1.sobs,data=toyr)
summary(lmfit)
plot(co2flux~sbs2009_1.sobs,data=toyr)
abline(lmfit)

temporary_soil<-my2009soilenv[-9]
sbs2009soilenv<-na.omit(as.matrix(temporary_soil[8:11]))
mycolnames<-colnames(sbs2009plants.sortedYtt, do.NULL = TRUE, prefix = "col")
```
## 8) Looking at the effect of treatment on the change in soil respiration over the years- time series analysis 
```{r}
myCO2ts<-resp[with(resp,order(trt,year,month)),]
myCO2ts<-myCO2ts[-1,]
myC_CO2ts<-filter(myCO2ts, trt=="C") %>% group_by(year, month)
myCts<-(count(myC_CO2ts,wt=co2flux))
myH_CO2ts<-filter(myCO2ts, trt=="H")
myHN_CO2ts<-filter(myCO2ts, trt=="HN")
myN_CO2ts<-filter(myCO2ts, trt=="N")
#<- filter(resp, taxa=="Rodent") %>% filter(plot_id==18) %>% group_by(year, month, plot_id) %>% count(species_id)
#myCO2ts$season <- NA
#myCO2ts$season <- myCO2ts$month %in% c(6:10)
#myCO2ts$season <- ifelse(myCO2ts$season == TRUE, "rain", "norain")
myCO2ts<-group_by(myCO2ts, year, month)
tester<-count(myCO2ts,wt=co2flux)

abund$nn <- abund$nn * 4
C_ts <- ts(abund$nn, frequency = 2, start = c(1977, 2))
plot.ts(abund.ts, type = "l", ylab = "Rodent Abundance (#/hectare)", xlab = "Time (year)", las = 1, ylim = c(0, 500))
#2
abund.sm <- SMA(abund$nn, n = 2)#used 2 bc we have condensed to hexamonthly datapoints
plot(abund.sm, type = "l", col = "#56e4c6", ylab = "Rodent Abundance (#/hectare)",
xlab = "Sample", las = 1, ylim = c(0, 500))
lines(abund$nn, col = "black")
legend(0, 475, col = c("#56e4c6", "black"), lty = c(1,1),
c("smooth", "non-smooth"), bty = "n", cex = 1)



abund.comp <- decompose(abund.ts)
# decomposition categories
plot(abund.comp)
#3. Test whether your data meet the assumption of stationarity.
adf.raw <- adf.test(abund.ts, alternative = "stationary")
print(adf.raw$p.value)
#for plot #18, the ts is NOT stationary: P=.77

#4. If it does not meet this assumption, explore ways to make your data stationary.
abund.ts.diff <- diff(abund.ts,lag=1)
adf.diff <- adf.test(abund.ts.diff, alternative = "stationary")
adf.diff$p.value
#now P=0.01, so null hypot of nonstationarity is rejected
plot.ts(abund.ts.diff, type = "l", ylab = "Rodent Abundance (#/hectare)", xlab = "Time (year)", las = 1)
#5. Examine and plot time lags using the partial autocorrelation function (PACF) and autocorrelation function (ACR).
acf(abund.ts.diff)
acf(abund.ts)
pacf(abund.ts.diff)
pacf(abund.ts)

#6. Use the tools outlined in the Handout to create an ARMA model.
fit18<-auto.arima(abund.ts,ic="aic",trace=T)
fit18<-arima(abund.ts,c(1,0,0),seasonal=list(order=c(2,1,0),period=2),include.mean=T)
#model is order (1,0,0) with nonzero mean
tsdiag(fit18)
pred18 <- predict(fit18, n.ahead = 20)
ts.plot(abund.ts, pred18$pred, lty = c(1,3))


```
## 9) Looking at the effect of treatment on the change in N mineralization over the years- time series analysis 
```{r}

```
## 10) TESTING PRACTICE WITH CATTERPLOT
```{r}
library(CatterPlots)
x <- -10:10
y <- -x^2 + 10
purr <- catplot(xs=x, ys=y, cat=3, catcolor=c(0,1,1,1))
cats(purr, -x, -y, cat=4, catcolor=c(1,0,1,1))

# for more fun ...
meow <- multicat(xs=x, ys=y, cat=c(1,2,3), catcolor=list(c(1,1,0,1),c(0,1,1,1)), canvas=c(-0.1,1.1, -0.1, 1.1))
morecats(purr, x, 10*sin(x)+40, size=0.05, cat=c(4,5,6), catcolor=list(c(0,0,1,1),c(0,1,0,1)), type="line")

# random cats
meow <- multicat(xs=x, ys=rnorm(21),
                 cat=c(1,2,3,4,5,6,7,8,9,10,11),
                 catcolor=list(c(1,0,0,1)),
                 canvas=c(-0.1,1.1, -0.1, 1.1),
                 xlab="some cats", ylab="other cats", main="Random Cats")
```

